---
title: "Homework 1"
subtitle: "4375 Machine Learning with Dr. Mazidi"
author: "Kinsey Mellon"
date: "6/05/22"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

This homework has two parts:

* Part 1 uses R for data exploration
* Part 2 uses C++ for data exploration

---

This homework is worth 100 points, 50 points each for Part 1 and Part 2. 

---

# Part 1: RStudio Data Exploration

**Instructions:** Follow the instructions for the 10 parts below. If the step asks you to make an observation or comment, write your answer in the white space above the gray code box for that step. 

## Step 1: Load and explore the data

* load library MASS (install at console, not in code)
* load the Boston dataframe using data(Boston)
* use str() on the data
* type ?Boston at the console
* Write 2-3 sentences about the data set below

Your commentary here:
This data set is the housing values in the suburbs of Boston. It has 506 rows and 14 columns. It includes things like crime rate per capita, the proportion of residential land zoned for lots over 25,000 sq.ft., the avg number of rooms per dwelling and other things.

```{r}
library(MASS)
data(Boston)
str(Boston)

```

## Step 2: More data exploration

Use R commands to:

* display the first few rows
* display the last two rows
* display row 5
* display the first few rows of column 1 by combining head() and using indexing
* display the column names

```{r}
head(Boston)
tail(Boston, n=2)
Boston[5,]
head(Boston[,1])
names(Boston)
```

## Step 3: More data exploration

For the crime column, show:

* the mean
* the median
* the range

```{r}
mean(Boston$crim)
median(Boston$crim)
range(Boston$crim)
diff(range(Boston$crim))
```

## Step 4: Data visualization

Create a histogram of the crime column, with an appropriate main heading. In the space below, state your conclusions about the crime variable:

Your commentary here:
The crime rate is generally pretty low. It's mostly below 10%


```{r}
hist(Boston$crim, main="Per Capita Crime Rate by Town")

```

## Step 5: Finding correlations

Use the cor() function to see if there is a correlation between crime and median home value. In the space below, write a sentence or two on what this value might mean. Also write about whether or not the crime column might be useful to predict median home value.

Your commentary here:
The value is negative so we can tell that crime rate and median home value are negatively correlated but overall the value of -0.388 is on the closer side to 0 so they're weakly correlated. I don't think the crime column is very useful to predict mean home value because they aren't strongly correlated.

```{r}
cor(Boston$crim, Boston$medv)
```

## Step 6: Finding potential correlations

Create a plot showing the median value on the y axis and number of rooms on the x axis. Create appropriate main, x and y labels, change the point color and style. [Reference for plots(http://www.statmethods.net/advgraphs/parameters.html) 

Use the cor() function to quantify the correlation between these two variables. Write a sentence or two summarizing what the graph and correlation tell you about these 2 variables.

Your commentary here:
The correlation tells us that the two variables have a strong positive correlation. The graph reflects that number by showing an increasing slope going to the right.

```{r}
plot(Boston$rm, Boston$medv, pch = 18, col = "blue",main = "Num of rooms and median home value", xlab = "Number of rooms in dwelling", ylab = "Median Home Value")
cor(Boston$rm, Boston$medv)
```

## Step 7: Evaluating potential predictors

Use R functions to determine if variable chas is a factor. Plot median value on the y axis and chas on the x axis. Make chas a factor and plot again.

Comment on the difference in meaning of the two graphs. Look back the description of the Boston data set you got with the ?Boston command to interpret the meaning of 0 and 1.  

Your commentary here:
When chas was made a factor the x axis changed from being 0 to 1, to simply 0 or 1. This was because when chas was changed to a factor its encodes it as qualitative data (either 1 or 0) instead of as integers plotting it on a range from 0 to 1.

```{r}
plot(Boston$chas, Boston$medv)
plot(as.factor(Boston$chas), Boston$medv)

```

## Step 8: Evaluating potential predictors

Explore the rad variable. What kind of variable is rad? What information do you get about this variable with the summary() function? Does the unique() function give you additional information? Use the sum() function to determine how many neighborhoods have rad equal to 24. Use R code to determine what percentage this is of the neighborhoods.

Your commentary here:
rad is the index of accessibility to radial highways. The summary() function gives the min, the 1st quartile, median, mean, 3rd quartile, and max. The unique() function returns the different values that the rad column holds, excluding any repeats of the same value.

```{r}
summary(Boston$rad)
unique(Boston$rad)
sum(Boston$rad == 24)
sum(Boston$rad == 24) / nrow(Boston) * 100

```

## Step 9: Adding a new potential predictor

Create a new variable called “far” using the ifelse() function that is TRUE if rad is 24 and FALSE otherwise. Make the variable a factor. Plot far and medv. What does the graph tell you? 

Your commentary here:
The graph shows us that houses that don't have a rad equal to 24 have a better median house value

```{r}
far <- ifelse(Boston$rad == 24, TRUE, FALSE)
plot(as.factor(far), Boston$medv)

```

## Step 10: Data exploration

* Create a summary of Boston just for columns 1, 6, 13 and 14 (crim, rm, lstat, medv)
* Use the which.max() function to find the neighborhood with the highest median value. See p. 176 in the pdf
* Display that row from the data set, but only columns 1, 6, 13 and 14
* Write a few sentences comparing this neighborhood and the city as a whole in terms of: crime, number of rooms, lower economic percent, median value.

Your commentary here:
The highest median value recorded was 50 which is considerable more than the mean of all the median values which is around 22. That neighborhood has a lower crime rate and lower economic percent compared to the mean of the whole city. On the other hand this neighborhood has a lower average number of rooms per dwelling.

```{r}
summary(Boston$crim)
summary(Boston$rm)
summary(Boston$lstat)
summary(Boston$medv)

i <- which.max(Boston$medv)
print(Boston[i, c(1,6,13,14)])
```

# Part 2: C++

In this course we will get some experience writing machine learning algorithms from scratch in C++, and comparing performance to R. Part 2 of Homework 1 is designed to lay the foundation for writing custom machine learning algorithms in C++. 
	
To complete Part 2, first you will read in the Boston.csv file which just contains columns rm and medv. 
	
--- 

In the C++ IDE of your choice:
	
1	Read the csv file (now reduced to 2 columns) into 2 vectors of the appropriate type. See the reading in cpp picture in Piazza. 

2	Write the following functions: 

*	a function to find the sum of a numeric vector

*	a function to find the mean of a numeric vector

*	a function to find the median of a numeric vector

*	a function to find the range of a numeric vector

*	a function to compute covariance between rm and medv (see formula on p. 74 of pdf)

*	a function to compute correlation between rm and medv (see formula on p. 74 of pdf); Hint: sigma of a vector can be calculated as the square root of variance(v, v)

3	Call the functions described in a-d for rm and for medv. Call the covariance and correlation functions. Print results for each function.







